---
title: "A UN report on GDP life Expectancy and CO2"
author: "Yesha"
date: "2023-01-18"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r load-packages, include=FALSE}
library (tidyverse)
library (knitr)

```
This report was prepared to the attention of the United Nations. It analyzes the relationship between a country's GDP, its life expectancy and CO2 emissions. Our goal is to determine to what degree a country’s economic strength or weakness may be related to its public health status and impact on climate pollution. We hypothesize that both life expectancy and CO2 emissions increase with a country's GDP.

```{r read-data, include=FALSE}
gapminder_1997 <- read_csv("data/gapminder_1997.csv")
```


```{r gdp-lifeExp-1997-plot, echo=FALSE, fig.width=3, fig.height=2, fig.align='center'}
# make plot of GDP vs LifeEXp

ggplot (data = gapminder_1997) +
aes(x = gdpPercap, y = lifeExp, color = continent) +
  labs ( x= "GDP per Capital", y= "Life Expectancy") +
geom_point()+
  scale_color_brewer(palette = "Set1")

```
```{r basic-country-stats, echo=FALSE}
# Number of countries in data
nCountries <- 
  gapminder_1997 %>%
  select (country) %>%
  n_distinct()

# minimum GDP
minGDP <- gapminder_1997 %>%
  summarize (round(min(gdpPercap)))%>%
pull()

maxGDP <- gapminder_1997 %>%
  summarize (round(max(gdpPercap)))%>%
  pull()

```
The number of countries in the data is `r nCountries`. The minimum GDP is `r minGDP`. The maximum GDP is `r maxGDP`.

```{r table}
gapminder_1997 %>%
  filter (continent == "Oceania") %>%
  kable()

```


